Embracing the Laws of Physics: \\ Reversible Models of Computation

### 2017-12-21 survey/survey.tex
continue with Data II.

\newtheorem{defn}{Definition}[section]

@@

This correspondance is rather fruitful. As logical expressions form
a semiring, we would expect that types too form a semiring. And indeed
they do -- at least up to \emph{type equivalence}.  The natural
numbers $\Nat$ are another semiring; it will turn out that, even though
the Curry-Howard correspondance has been extremely fruitful for
programming language research, it is the semiring $\Nat$ which
will be a better model for finite structured types.

>>>>

\newtheorem{defn}{Definition}[section]
\newtheorem{prop}{Proposition}[section]

@@

This correspondance is rather fruitful. As logical expressions form
a commutative semiring, we would expect that types too form a commutative
semiring. And indeed they do -- at least up to \emph{type equivalence}.  The
natural numbers $\Nat$ are another commutative semiring; it will turn out that,
even though the Curry-Howard correspondance has been extremely fruitful for
programming language research, it is $\Nat$ which will be a better model for
finite structured types.

\begin{definition}
  A \emph{commutative semiring} (sometimes called a \emph{commutative
    rig} (commutative ri\emph{n}g without negative elements) consists of a
  set $R$, two distinguished elements of $R$ named 0 and 1, and two
  binary operations~$+$ and $\cdot$, satisfying the following
  relations for any $a,b,c \in R$:
\[\begin{array}{rcl}
0 + a &=& a \\
a + b &=& b + a \\
a + (b + c) &=& (a + b) + c \\
\\
1 \cdot a &=& a \\
a \cdot b &=& b \cdot a \\
a \cdot (b \cdot c) &=& (a \cdot b) \cdot c \\
\\
0 \cdot a &=& 0 \\
(a + b) \cdot c &=& (a \cdot c) + (b \cdot c)
\end{array}\]
\end{definition}

\begin{prop}
The structure $\left(\mathbb{B}, \false, \true, \lor, \land\right)$
is a commutative semiring.
\end{prop}

We cannot immediately use the above definition in a typed setting.
First, types do not naturally want to be put together into a ``set''.
This can be fixed if we replace the set $R$ with a universe $U$, and
rather than insisting that we have $0 \in R$, we have a typing
judgement $\bot : U$ (and similarly for the other items).

Our first instinct would be to similarly replace $=$ with
$\equiv$, i.e. propositional equality. This bring the second
problem to the fore: this is false. For example, unless $A \equiv B$,
$A \times B \equiv B \times A$ is not normally provable%
\footnote{Not without assuming that $U$ is univalent, which we
don't.}. But it should be clear that $A \times B$ and $B \times A$
contain equivalent information. In other words, we would like to
be able to witness that $A \times B$ can be reversibly deformed
into $B \times A$. Which motivates the introduction of 
type \emph{equivalences}. To do this, we need a few important
auxilliary concepts first.

\begin{definition}[homotopy]
\label{def:homotopy}
Two functions $f,g:A \rightarrow B$ are \emph{homotopic} if
$\forall x:A. f(x) = g(x)$. We denote this $f \sim g$.
\end{definition}

\noindent It is easy to prove that homotopies (for any given function
space $A \rightarrow B$) are an equivalence relation.

\begin{definition}[Quasi-inverse]
\label{def:quasi}
For a function $f : A \rightarrow B$, a \emph{quasi-inverse} is a
triple $(g, \alpha, \beta)$, consisting of a function
$g : B \rightarrow A$ and two homotopies
$\alpha : f \circ g \sim \mathrm{id}_B$ and
$\beta : g \circ f \sim \mathrm{id}_A$.
\end{definition}

\begin{definition}[Equivalence of types]
  Two types $A$ and $B$ are equivalent $A ≃ B$ if there exists a
  function $f : A \rightarrow B$ together with a quasi-inverse for $f$.
\end{definition}

\jc{talk about the definition without quasi-inverse? Continue with
types being a commutative semiring in the above sense.}
